CS~2510: Course Charter \& Syllabus

### 2012-05-25 admin/charter/CS2510.tex
Revised charter for 2510.

\paragraph{Course Description:}
%
Introduces the fundamentals of the design of class hierarchies in object-oriented programming with emphasis on the use of abstractions to design reusable code. The course also covers basic reasoning techniques concerning the algorithmic behavior of functions and methods, especially for linear structures (e.g., lists, stacks, queues) and value-dependent structures (e.g., binary search trees, hash tables).

\paragraph{Prerequisites:}
%
CS~2110. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2500 or concurrently.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Class Hierarchies}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.

\paragraph{Topics Covered:}
%
Classes, class containment, class hierarchies: data definitions and methods.
%
Designing abstractions and reusable code: interfaces, abstract classes, subclasses, function objects, abstracting over traversals, polymorphism, abstract data type, algorithms parametrized over the data type and functional behavior. 
%
Equality and unit test design.
%
Introduction to algorithm complexity illustrated on linear structures and their algorithms: lists, queues, stacks; direct access structures and their algorithms: sorting and searching, value-dependent (parametric) data structures and related algorithms: binary search trees, heaps, hash tables.
%
Composite structures and algorithms: graphs, graph traversals and searches, backtracking.
%
Program re-use: use of existing libraries.
%
Designing user interactions: the Model-View-Control architecture.
%
Exposure to design patterns: Chain of Responsibility, Template and Hook, Visitor, Iterator, Command, State, Adapter

@@

\paragraph{Course Objectives:}
%
The goal is to help the student understand the principles of class based program design using an object-oriented programming language(s)---not just Java. Java-like syntax is used early, with the transition to standard Java later in the course, to illustrate how the program design principles are used in practical applications. Using a concrete language also presents an opportunity to discuss the strengths and weaknesses of languages and language paradigms.

>>>>

\paragraph{Course Description:}
%

Introduces the fundamentals of the design of class hierarchies in
object-oriented programming with emphasis on representation-dependent
program design with objects. The course also covers basic reasoning
techniques concerning the correctness and algorithmic behavior of
computations.

\paragraph{Prerequisites:}
%
CS~2500. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2510 or concurrently.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Classes}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.

\paragraph{Topics Covered:}
%
Objects, classes, class hierarchies.  Data definitions and methods.
%
Class diagrams.
%
Interfaces, abstract classes, subclasses.  Types.  Polymorphism.
%
Accumulators, iterators, loops.
%
Function objects, comparators, visitors.  Sorting: quick sort, merge
sort, heap sort.  Notions of equality.
%
Java collections library.
%
Overriding {\tt equals} and {\tt hashCode}.
%
Testing: unit testing, stress testing, and property-based random testing.
%
Binary trees, binary search, heaps, hash tables.
%
Invariants.
%
Graphs, graph traversal, generative recursion.  Depth-first,
breadth-first, and priority-queue based search.
%
Design patterns.
%
Mutation and stateful programming.
%
Indexed data structures.

@@

\paragraph{Course Objectives:}
%
The goal is to help the student understand the principles of class
based program design using an object-oriented programming
language. Java is used to illustrate how the program design principles
are used in practical applications. Using a concrete language also
presents an opportunity to discuss the strengths and weaknesses of
languages and language paradigms.
